---
title: "My Quarto Practice 2"
author: "Bella Kintigh"
format: gfm
prefer-html: true
knitr: 
  opts_chunk: 
    fig.path: "../Output/"
theme: lux
---

# Introduction 

Today we are learning how to work with figures in Quarto for OCN 682.

## Load the libraries

```{r}
#| message: false
#| warning: false

library(tidyverse)
library(here)
library(palmerpenguins)
library(kableExtra)
```

# Make a plot

```{r}
#| message: false
#| warning: false
#| out-width: "70%"
#| fig-align: "center"
#| label: fig-penguin
#| fig-cap: "This figure shows length by bill depth"

library(palmerpenguins)
library(tidyverse)

penguins %>%
  ggplot(aes(x = bill_length_mm, 
             y = bill_depth_mm, 
             color = species)) +
  geom_point()
```

When you look at @fig-penguin you can see there is a positive relationship between bill length and bill depth 

# Practice Table

| Time  | Session | Topic    |
|:------|:-------:|---------:|
| _left_ | _center_| _right_ |
| 01:00  | 1       | Anatomy |
| 01:50  |         | *Break* |
| 02:00  | 2       | Tables  |
| 02:45  |         | *Break* |

# Table with kable

```{r}
#| message: false
#| warning: false
#| label: tbl-penguin
#| tbl-cap: "My awesome penguin table"

library(kableExtra)
library(palmerpenguins)
library(tidyverse)

penguins %>%
  group_by(species) %>%
  summarise(billmean = mean(bill_length_mm, na.rm = TRUE)) %>%
  kbl() # make it a kable table
```

Table @tbl-penguin has information on the mean bill length by species.

# Add Some Flair to the Table

```{r}
library(kableExtra)
library(palmerpenguins)
library(tidyverse)

penguins %>%
  group_by(species) %>%
  summarise(billmean = round(mean(bill_length_mm, na.rm = TRUE))) %>%
  kbl()  %>% # make it a kable table
  kable_classic() %>% # add a theme
  row_spec(2, bold = TRUE, color = "white", background = "blue") %>% # highlight row 2
  kable_styling(full_width = FALSE) # don't make it so wide
```

