---
title: "Tidy Tuesday - 2025.11.11"
author: "Bella Kintigh"
format: html
warning: FALSE
message: FALSE
theme: minty
---

# Purpose

In this assignment, I aim to:

-   Practice tidying, filtering, and visualizing time-series and country-level data
-   Explore regional differences in TB incidence and mortality rates
-   Learn and apply at least one new data visualization or transformation technique
-   Reflect on how public health data can guide policy, equity, and resource allocation
-   My new skill to learn: interactive maps

# Libraries

```{r}
library(tidyverse)
library(tidytuesdayR)
library(leaflet)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(here)
```

# Data

This week’s TidyTuesday assignment explores global tuberculosis (TB) burden data compiled by the World Health Organization (WHO). The dataset includes TB incidence, mortality, and population estimates across multiple countries and years. This assignment offers an opportunity to examine patterns in TB outcomes, evaluate geographic or temporal disparities, and build skills in data visualization and interpretation of global health metrics.

## Load in data

```{r}
# TidyTuesday data
tuesdata <- tidytuesdayR::tt_load('2025-11-11')

# World country shapes
world <- rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")
```

## Data wrangling

```{r}
# Subset for only 2021
tb <- tuesdata$who_tb_data %>%
  filter(year == 2021) %>%
  select(country, inc = e_inc_100k)

# Join TB data to world map
map_data <- world %>%
  left_join(tb, by = c("name" = "country"))
```

# Build leaflet map 

```{r}
leaflet(map_data) %>%
  addTiles() %>% # Adds a default OpenStreetMap tile background
  addPolygons(
    fillColor = ~colorBin("YlOrRd", inc, bins = 5)(inc),
     # Shades each country based on TB incidence (`inc`) using a yellow-to-red color scale
     # `colorBin()` creates 5 color breaks
    weight = 0.5, # border around each country
    color = "#333", # border color
    fillOpacity = 0.7, # fill for the country is slightly transparent
    label = ~paste0(name, ": ", round(inc, 1), " cases per 100k"), # text that hovers over the country
    highlightOptions = highlightOptions(
      weight = 2, # thicker border when hovering
      color = "#000", # color of border (little darker)
      fillOpacity = 0.9, # slightly less transparent
      bringToFront = TRUE # brings hovered country to front
    )
  )
```

# Save a Snapshot of Interactive Map 

```{r}
# Use combined tb and map df

ggplot(map_data, aes(fill = inc)) +
  geom_sf(color = "white") +
  scale_fill_viridis_c(option = "C", na.value = "gray90") +
  labs(
    title = "TB Incidence by Country (2021)",
    fill = "Cases per 100k"
  ) +
  theme_minimal()

# Save it 

ggsave(here("Tidy_Tuesday", "11_11_2025", "Output", "tb_map_snapshot.png"), 
       width = 8,
       height = 5,
       dpi = 300)
```









